---
title: "RWork-sheet_Sabio#4b.Rmd"
output: pdf_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r 1}
vectorA<- c(1,2,3,4,5)
matrixA<- matrix( nrow = 5, ncol = 5)

  for (i in 1:5){
    for (j in 1:5){
      matrixA[i, j] <- abs(vectorA[i] - vectorA[j])
    }
  }
print(matrixA)
```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure 2.

```{r 2}
n <- 5
  for ( i in 1:n){
    for (j in 1:i){
    cat("*")
  }
cat("\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r 3}
get_int <- function(prompt) {
  repeat {
    input <- as.integer(readline(prompt))
    if (!is.na(input) && input > 0) {
      return(input)
    } else {
      cat("Please enter a valid positive integer.\n")
    }
  }
}

n <- get_int("Enter the starting number for the Fibonacci sequence: ")

a <- 0
b <- 1
 cat("Fibonacci sequence starting from", n, "and up to 500:\n")
repeat {
  if (a >= n) {
    cat(a, " ")
  }
  c <- a + b
  a <- b
  b <- c
  if (a > 500) {
    break
  }
}
```
4. Import the dataset as shown in Figure 1 you have created previously.
a.What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r 4a}
install.packages("readxl")
library(readxl)
data <- read_excel("Shoe Sizes.xlsx")
data 
head(data, n=6)
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r 4b}
maleSubset <- subset(data, Gender...3  == "M")
femaleSubset <- subset(data, Gender...3  == "F")

Male <- nrow(maleSubset)
Female <- nrow(femaleSubset)

cat("Number of observations of Male in Gender...3 is: ", Male, "\n")
cat("Number of observations of Female in Gender...3 is: ", Female, "\n")

maleSubset1 <- subset(data, Gender...6  == "M")
femaleSubset1 <- subset(data, Gender...6  == "F")

Male1 <- nrow(maleSubset1)
Female1 <- nrow(femaleSubset1)

cat("Number of observations of Male in Gender...6 is: ", Male1, "\n")
cat("Number of observations of Female in Gender...6 is: ", Female1, "\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r 4c}
gender <- c("Female","Male")
counts <- c(2, 1)
data <- data.frame(gender = gender, Count = counts)
bar_colors <- c("pink", "violet")
barplot(data$Count, names.arg = data$Gender, col = bar_colors,
        main = "Count of Categories",
        xlab = "Gender",
        ylab = "Count")
legend("topright",
       legend = data$gender,
       fill = bar_colors,
       border = "black")
```           