---
title: "RWork-sheet_Sabio#4b.Rmd"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
date: "2023-11-08"
---

```{r setup, include=FALSE}
```
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r 1}
vectorA<- c(1,2,3,4,5)
matrixA<- matrix( nrow = 5, ncol = 5)

  for (i in 1:5){
    for (j in 1:5){
      matrixA[i, j] <- abs(vectorA[i] - vectorA[j])
    }
  }
print(matrixA)
```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure 2.

```{r 2}
n <- 5
  for ( i in 1:n){
    for (j in 1:i){
    cat("*")
  }
cat("\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r 3}
fibSequence <- as.integer(readline(prompt = "Enter the number of terms: "))

a <- 0
b <- 1
cat("Fibonacci Sequence:", a, b)

repeat {
c <- a + b
if (c > 500) {
break
}
cat(", ",c)
a <- b
b <- c
}
```
4. Import the dataset as shown in Figure 1 you have created previously.
a.What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r 4a}
install.packages("readxl")
library(readxl)
data <- read_excel("Shoe Sizes.xlsx")
data 
head(data, n=6)
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r 4b}
maleSubset <- subset(data, Gender...3  == "M")
femaleSubset <- subset(data, Gender...3  == "F")

Male <- nrow(maleSubset)
Female <- nrow(femaleSubset)

cat("Number of observations of Male in Gender...3 is: ", Male, "\n")
cat("Number of observations of Female in Gender...3 is: ", Female, "\n")

maleSubset1 <- subset(data, Gender...6  == "M")
femaleSubset1 <- subset(data, Gender...6  == "F")

Male1 <- nrow(maleSubset1)
Female1 <- nrow(femaleSubset1)

cat("Number of observations of Male in Gender...6 is: ", Male1, "\n")
cat("Number of observations of Female in Gender...6 is: ", Female1, "\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r 4c}
count <- c(Male, Female)
gender <- c("Male", "Female")
barplot(count,
names.arg = gender,
main = "The number of Males and Females in Household Data",
xlab = "Gender",
ylab = "Count",
col = c("violet", "pink"))
legend("topright",
legend = gender,
fill = c("violet", "pink"))
```   
5a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r 5a}
expenses <- c(60, 10, 5, 25)
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")

percentages <- round((expenses / sum(expenses)) * 100, 1)

pie(expenses, main="Dela Cruz Family Monthly Expenses", col = rainbow(length(expenses)),
    labels = categories, cex = 0.8)
legend (1.4,0.5,
        c("Food","Electricity","Savings","Miscelleneous"), cex = 0.7, fill =rainbow(length(expenses)))
```
6a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r 6a}
data("iris")
str(iris)
#It is a data frame that has 150 observation of 5 variables which is sepal.length, sepal.width, petal.length, petal.width.
```
b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r 6b}
meanVal <- c(mean(iris$Sepal.Length), mean(iris$Sepal.Width), mean(iris$Petal.Length), mean(iris$Petal.Width))
meanVal
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r 6c}
pie(table(iris$Species),
main = "Species distribution",
labels = levels(iris$Species),
col = c("lightblue","lightgreen","pink"))
legend("topright", legend = levels(iris$Species), fill = c("lightblue","lightgreen","pink") , title = "Species")
```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r 6d}
subsetSetosa <- tail(subset(iris, Species == "setosa"), 6)
subsetSetosa
subsetVersicolor <- tail(subset(iris, Species == "versicolor"), 6)
subsetVersicolor
subsetVirginica <- tail(subset(iris, Species == "virginica"), 6)
subsetVirginica
```
e. Create a scatterplot of the sepal.length and sepal.width using the different species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.
```{r 6e}
iris$Species<- as.factor(iris$Species)
plot(iris$Sepal.Length,iris$Sepal.Width,pch = as.integer(iris$Species),col = as.integer(iris$Species), main = "Iris Dataset", sub = "Sepal Width and Length", xlab = "Sepal Length", ylab = "Sepal width")

as.factor(iris$Species)
## [1] setosa setosa setosa setosa setosa setosa
## [7] setosa setosa setosa setosa setosa setosa
## [13] setosa setosa setosa setosa setosa setosa
## [19] setosa setosa setosa setosa setosa setosa
## [25] setosa setosa setosa setosa setosa setosa
## [31] setosa setosa setosa setosa setosa setosa
## [37] setosa setosa setosa setosa setosa setosa
## [43] setosa setosa setosa setosa setosa setosa
## [49] setosa setosa versicolor versicolor versicolor versicolor
## [55] versicolor versicolor versicolor versicolor versicolor versicolor
## [61] versicolor versicolor versicolor versicolor versicolor versicolor
## [67] versicolor versicolor versicolor versicolor versicolor versicolor
## [73] versicolor versicolor versicolor versicolor versicolor versicolor
## [79] versicolor versicolor versicolor versicolor versicolor versicolor
## [85] versicolor versicolor versicolor versicolor versicolor versicolor
## [91] versicolor versicolor versicolor versicolor versicolor versicolor
## [97] versicolor versicolor versicolor versicolor virginica virginica
## [103] virginica virginica virginica virginica virginica virginica
## [109] virginica virginica virginica virginica virginica virginica
## [115] virginica virginica virginica virginica virginica virginica
## [121] virginica virginica virginica virginica virginica virginica
## [127] virginica virginica virginica virginica virginica virginica
## [133] virginica virginica virginica virginica virginica virginica
## [139] virginica virginica virginica virginica virginica virginica
## [145] virginica virginica virginica virginica virginica virginica
## Levels: setosa versicolor virginica
```
f. Interpret the result.
```{r 6f}
#the scatterplot result represents the relationship between sepal length and width for different species(setosa,versicolor,virginica)and each point corresponds to an observation in the dataset,differentiated by color and symbol based on species.
```
7a. Rename the white and black variants by using gsub() function.
```{r 7a}
install.packages("readxl")
library(readxl)
alexaFile <- read_excel("alexa_file.xlsx")

alexaVar <- gsub("Black Plus", "Black Plus", alexaFile$variation)
alexaFile$variation <- gsub("Black Show", "Black Show", alexaFile$variation)
alexaFile$variation <- gsub("Black Spot", "Black Spot", alexaFile$variation)
alexaFile$variation <- gsub("Black Dot", "Black Dot", alexaFile$variation)
alexaFile$variation <- gsub("White Dot", "White Dot", alexaFile$variation)
alexaFile$variation <- gsub("White Plus", "White Plus", alexaFile$variation)
alexaFile$variation <- gsub("White Show", "White Show", alexaFile$variation)
alexaFile$variation <- gsub("White Spot", "White Spot", alexaFile$variation)

```
7b.Get the total number of each variations and save it into another object. Save the object as variations.RData. Write the R scripts. What is its result?
```{r 7b}
library(dplyr)

save(alexaFile, file = "variations.RData")
load("variations.RData")
alexaVar <- alexaFile%>%count(alexaFile$variation)
alexaVar

```
7c.From the variations.RData, create a barplot(). Complete the details of the chartwhich include the title, color, labels of each bar.
```{r 7c}
barplot(
height = alexaVar$n,
names.arg = alexaVar$`alexaFile$variation`,
col = "pink",
main = "Alexa Varations",
las = 2,
cex.names = 0.58
)

```

7d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by side. Complete the details of the chart.
```{r 7d}
par(mfrow = c(1, 2))
blackVariants <- alexaVar[1:5,]
whiteVariants <- alexaVar[12:16,]
barplot(
height = blackVariants$n,
names.arg = blackVariants$`alexaFile$variation`,
main = "Black Variants",
col = rainbow(5),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)
barplot(
height = whiteVariants$n,
names.arg = whiteVariants$`alexaFile$variation`,
main = "White Variants",
col = rainbow(5),
xlab = 'Total Numbers',
ylab = 'Frequency',
cex.names = 0.35,
)
```