---
title: "RWorksheet_Sabio#4c.Rmd."
output: pdf_document
date: "2023-11-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1a. Show your solutions on how to import a csv file into the environment.
```{r 1}
mpgData <- read.csv("mpg.csv")
mpgData
```
b. Which variables from mpg dataset are categorical?
```{r 1b}
library(ggplot2)
data(mpg)
str(mpg)
```
c. Which are continuous variables?
In the mpg dataset, numeric variables like engine displacement ('displ'), miles per gallon on the highway ('hwy') and in the city ('cty'), as well as the variable 'year', represent quantifiable aspects. However, it's important to recognize that not all numeric variables are strictly continuous; for example, 'year' might signify discrete categories or ordered values rather than constituting a continuous numerical range.

2.1 Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r 2}
mostManufac <- names(sort(table(mpg$manufacturer), decreasing = TRUE))[1]
mostManufac
mostVar <- names(sort(table(mpg$model), decreasing = TRUE))[1]
mostVar
```
2a.Group the manufacturers and find the unique models. Show your codes and result.
```{r 2a}
library(dplyr)
uniqueModels <- mpg %>%
  group_by(manufacturer) %>%
  summarize(unique_models = n_distinct(model))
uniqueModels
```
2b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r 2b}
library(ggplot2)
ggplot(uniqueModels, aes(x = manufacturer, y = unique_models)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  labs(x = "Manufacturer", y = "Unique Models",
       title = "Unique Models by Manufacturer") +
  theme_minimal()
```
2.2a Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
```{r 2.2a}
library(ggplot2)
relationshipMM <- ggplot(mpg, aes(model, manufacturer)) +
  geom_point() +
  labs(x = "Model", y = "Manufacturer",
       title = "Relationship between Model and Manufacturer")
relationshipMM
```
#The plot has the potential to illustrate the distribution of car models across different manufacturers.

b. For you, is it useful? If not, how could you modify the data to make it more informative?
```
I find the plot less useful due to potential overlapping points that hinder the clarity of the relationship between car models and manufacturers. I would consider grouping or combining data and experimenting with alternative geometric graph representations to improve visual clarity and comprehension.
```
3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r 3}
library(ggplot2)
top20_observations<- head(mpg, 20)
ggplot(top20_observations, aes(x = year, y = model)) +
  geom_point() +
  labs(x = "Year", y = "Model", title = "Top 20 Observations: Model vs Year")
```
4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r 4}
library(dplyr)

cars_per_model <- mpg %>%
  group_by(model) %>%
  summarize(num_cars = n())
cars_per_model
```
4a.Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels and colors. Show code and results.
```{r 4a}
library(ggplot2)

modelCounts <- mpg %>%
  group_by(model) %>%
  summarise(Num_Cars = n()) %>%
  top_n(20, Num_Cars)

ggplot(modelCounts, aes(x = Num_Cars, y = reorder(model, Num_Cars), fill = model)) +
 geom_bar(stat = "identity") +
 scale_fill_manual(values = rainbow(nrow(modelCounts))) +
 labs(x = "Number of Cars", y = "Model", title = "Top 20 Number of Cars per Model") +
 theme(axis.text.y = element_text(angle = 0))
```
4b.Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r 4b}
top20 <- head(mpg, 20)
ggplot(top20, aes(x = model, fill = model)) +
  geom_bar() +
  labs(x = "Count", y = "Model", fill = "Model",
       title = "Top 20 Car Models Count") +
  coord_flip() + 
  theme(axis.text.y = element_text(hjust = 1)) + 
  scale_fill_brewer(palette = "Set3")
```
5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.
```{r 5}
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement")
```
5a. How would you describe its relationship? Show the codes and its result.
```
#The scatter plot illustrates the relationship between the number of cylinders (cyl) and engine displacement (displ), with the color of each point representing engine displacement. 
```
6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r 6}
library(ggplot2)
data(mpg)
mpg$year <- as.factor(mpg$year)

ggplot(mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  labs(x = "Engine Displacement", y = "Highway MPG",
       title = "Relationship between Displacement and Highway MPG\nMapped with Year") +
  scale_color_brewer(palette = "RdYlBu")
```
6. Import the traffic.csv onto your R environment.
```{r 6.1}
trafficData <- read.csv("traffic.csv")
trafficData
```
6a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.
```{r 6a}
dim(trafficData)
str(trafficData)
```
6b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r 6b}
subset_junctions=subset(trafficData, Junction < 10)
subset_junctions
```
6c. Plot each junction in a using geom_line(). Show your solution and output.
```{r 6c}
library(ggplot2)
ggplot(subset_junctions, aes(x = Junction, y = Vehicles, group = Junction, color = Junction)) +
  geom_line() +
  labs(x = "Junction", y = "Vehicle", title = "Traffic Flow at Different Junctions") +
  theme_minimal()

```
7.From alexa_file.xlsx, import it to your environment.
```{r 7}
library(readxl)
alexaData <- read_excel("alexa_file.xlsx")
alexaData
```
a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.
```{r 7a}
dim(alexaData)
str(alexaData)
```
b. Group the variations and get the total of each variations. Use dplyr package. Show solution and answer.
```{r 7b}
library(dplyr)

alexaVar <- alexaData %>%
  group_by(variation) %>%
  summarise(Total_Count = n())
alexaVar
```
c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.
```{r 7c}
library(ggplot2)
ggplot(alexaData, aes(x = variation)) +
  geom_bar(fill = "pink") +
  labs(x = "Variations", y = "Frequency", title = "Frequency of Variations in Alexa Dataset") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.
```{r 7d}
library(ggplot2)
library(dplyr)

review_counts <- alexaData %>%
  group_by(date) %>%
  summarise(Num_Reviews = n())
ggplot(review_counts, aes(x = date, y = Num_Reviews, group = 1)) +
  geom_line(color = "red") +
  labs(x = "Date", y = "Number of Verified Reviews", title = "Verified Reviews Over Time") +
  theme_minimal()
```
e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```{r 7e}
library(dplyr)
library(ggplot2)

variation_ratings <- alexaData %>%
  group_by(variation) %>%
  summarise(Avg_Rating = mean(rating, na.rm = TRUE)) %>%
  arrange(desc(Avg_Rating))

variation_ratings$variation[which.max(variation_ratings$Avg_Rating)]

ggplot(variation_ratings, aes(x = reorder(variation, Avg_Rating), y = Avg_Rating)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Variations", y = "Average Rating", title = "Relationship of Variations and Ratings") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```