---
title: "RWorksheet_5.Rmd"
output:
  pdf_document: default
  word_document: default
date: "2023-12-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Create a data frame for the table below. Show your solution.
a. Compute the descriptive statistics using different packages (Hmisc and pastecs).
Write the codes and its result.
```{r 1}
library(Hmisc)
library(pastecs)
studentData<- data.frame(
  Student <- c(1:10),
  preTest <- c(55,54,47,57,51,61,57,54,63,58),
  postTest <- c(61,60,56,63,56,63,59,56,62,61))

colnames(studentData) <- c("Student", "preTest","postTest")
studentData
```
2. The Department of Agriculture was studying the effects of several levels of a fertilizer on the growth of a plant. For some analyses, it might be useful to convert the fertilizer levels to an ordered factor.
```{r 2}
fertilizerData<- c(10,10,10, 20,20,50,10,20,10,50,20,50,20,10.)
ordered(fertilizerData)
```
3. Abdul Hassan, president of Floor Coverings Unlimited, has asked you to study the exercise levels undertaken by 10 subjects were “l”, “n”, “n”, “i”, “l” , “l”, “n”, “n”, “i”, “l” ; n=none, l=light, i=intense.
```{r 3}
exerciseLevels<- c("l", "n", "n", "i", "l", "l", "n", "n", "i", "l")
factor_exerciselevels<- factor(exerciseLevels)
levels(factor_exerciselevels) <- c("none","light","intense")
factor_exerciselevels
```
4a. Apply the factor function and factor level. Describe the results.
```{r 4}
state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld",
"vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt",
"wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw",
"vic", "vic", "act")
factor_state<-factor(state)
levels(factor_state) 
factor_state
```
5. From #4 - continuation:
• Suppose we have the incomes of the same tax accountants in another vector (in suitably large units of money).
```{r 5}
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54,
             62, 69, 70, 42, 56, 61, 61, 61, 58, 51, 48,
             65, 49, 49, 41, 48, 52, 46, 59, 46, 58, 43)

data <- data.frame(State = factor_state, Income = incomes)
print(data)
```
a. Calculate the sample mean income for each state we can now use the special function tapply():
```{r 5a}
mean_income <- tapply(incomes, factor_state, mean)
print(mean_income)
```
b. Copy the results and interpret.
 act      nsw       nt      qld       sa      tas 
44.50000 57.33333 55.50000 53.60000 55.00000 60.50000 
     vic       wa 
56.00000 52.25000 

Interpretation:
These values represent the average income for tax accountants from all the states and territories of Australia.

6.Calculate the standard errors of the state income means (refer again to number 3)
```{r 6}
stdError <- function(x) sqrt(var(x) / length(x))
errors_means <- tapply(incomes, factor_state, stdError)

print(errors_means)
```
6.
a. What is the standard error? Write the codes.
b. Interpret the result.

Interpretation:
The overall standard error serves as an indicator of how precise the sample mean estimate is for the entire group of tax accountants.

```{r 6a}
total_stdError <- stdError(incomes)
total_stdError
```
7. Use the titanic dataset.
a. subset the titatic dataset of those who survived and not survived. Show the codes and its result.
```{r 7}
library(titanic)

data("titanic_train")
titanicData <- titanic_train

people_survived <- subset(titanicData, Survived == 1)
people_not_survived <- subset(titanicData, Survived == 0)

head(people_survived)
head(people_not_survived)
```
8. The data sets are about the breast cancer Wisconsin. The samples arrive periodically as Dr. Wolberg reports his clinical cases.

a. describe what is the dataset all about.
< The Breast Cancer Wisconsin dataset consists of various quantitative and qualitative features extracted from biopsy samples, each associated with a unique ID number. These features are clump thickness, size uniformity,shape uniformity,marginal adhesion,epithelial size,bare nucleoli,bland chromatin,normal nucleoli,mitoses,class.

d. Compute the descriptive statistics using different packages. Find the values of:
d.1 Standard error of the mean for clump thickness.
d.2 Coefficient of variability for Marginal Adhesion.
d.3 Number of null values of Bare Nuclei.
d.4 Mean and standard deviation for Bland Chromatin
d.5 Confidence interval of the mean for Uniformity of Cell Shape.

```{r 8d}
breastCancer_Data <- read.csv("breastcancer_wisconsin.csv")
breastCancer_Data
```
```{r 8d.1}
#8.d1
stdError_clumpThickness <- sd(breastCancer_Data$clump_thickness) / sqrt(length(breastCancer_Data$clump_thickness))

stdError_clumpThickness

#8.d2
coefficient_marginalAdhesion <- sd(breastCancer_Data$marginal_adhesion) / mean(breastCancer_Data$marginal_adhesion)

coefficient_marginalAdhesion

#8.d3
nullValues_bareNuclei <- sum(is.na(breastCancer_Data$bare_nuclei))

nullValues_bareNuclei

#8.d4
mean_blandChromatin <- mean(breastCancer_Data$bland_chromatin)
sd_blandChromatin <- sd(breastCancer_Data$bland_chromatin)

mean_blandChromatin
sd_blandChromatin

#8.d5
confidenceInterval_uniformityCellShape <- t.test(breastCancer_Data$shape_uniformity)$conf.int

confidenceInterval_uniformityCellShape

```
8d. How many attributes?
```{r 8d2}
attributes(breastCancer_Data)
```
8e. Find the percentage of respondents who are malignant. Interpret the results.
```{r 8e}
malignantNum <- sum(breastCancer_Data$malignant == 1)

total_respondents <- nrow(breastCancer_Data)

malignant_prcnt <- (malignantNum / total_respondents) * 100
cat("Percentage of respondents who are malignant:", malignant_prcnt, "%\n")

```
9.Export the data abalone to the Microsoft excel file. Copy the codes.
```{r 9}
library(xlsx)
abalone<- read.csv("abalone.csv")
xlsx::write.xlsx(abalone, "abalone.xls", col.names=TRUE, row.names=TRUE, sheetName="sample")

library("AppliedPredictiveModeling")
head(abalone)
summary(abalone)
```